---
title: "Data Analysis report"
output: html_document
---

This is a report for data analysis assignement given for the reproducible reseatch course.
Begin by loading the data to be used. First unzip the data then load it:
```{r chunk1:unzip data}
setwd("~/Coursera/RepData_PeerAssessment1")
unzip("~/Coursera/RepData_PeerAssessment1/activity.zip")
activity <- read.csv("activity.csv")

```


View the data

```{r chunk2}
str(activity)
```

## What is the total number of steps taken per day?

```{r, echo=FALSE}
daily.steps<-with(na.omit(activity),aggregate(steps,by=list(date),sum))
colnames(daily.steps) <-c("date","steps")
daily.steps[1:8,]
```
The following histogram shows the total number of steps raken each day.
```{r}
library(ggplot2)
ggplot(daily.steps, aes(x=steps)) +
    geom_histogram(fill="brown",binwidth=800) +
    labs(title="Histogram of daily steps",
         x = "Number of steps per day",y="Number of times in a day") + theme_bw()
```

```{r, echo=FALSE}
mns1<-with(daily.steps, mean(steps,na.rm=TRUE))
mdn1<-with(daily.steps, median(steps,na.rm=TRUE))

```

The mean total number of steps taken  per day is `r mns1`

The median total number of steps taken  per day is `r mdn1`


## What is the average daily activity pattern?
```{r, echo=FALSE}
#Aggregate the steps per interval.
i.activity <- with(na.omit(activity),aggregate(steps,by=list(interval),mean))
colnames(i.activity)<- c("interval","steps")
#View the first ones
i.activity[1:6,]
#Which interval had the maxiumum daily steps?
max.i <-i.activity[,1][which(i.activity[,2]==max(i.activity[,2]))]
```
Averaged daily steps of the 5-minute interval time series plot.
``` {r, echo=FALSE}
with(i.activity, plot(interval,steps,type="l",col="red",lty=2,ylab="Average daily steps",xlab="5-minute interval"))
```

The `r max.i`-th 5-min interval contains the maximum number of steps.

## Imputing missing values
First compute the number of rows with steps coded as NA
```{r, echo=TRUE}
#number of rows with NA steps
act.na<-nrow(subset(activity,activity$steps!=is.na(activity$steps))) 
```
There are `r act.na` rows with NA steps values.

##Imputation strategy
I will use median because there are so many zeroes for that interval to fill in the missing values and keeping it simple

```{r}
uint <- unique(activity$interval)
activity.imp <-NULL
for (k in 1:length(uint)){
activity.u <- subset(activity, activity$interval==uint[k])
before<-activity.u$steps[1:20]#before impuatation
activity.u$steps<-ifelse(is.na(activity.u$steps),median(na.omit(activity.u$steps)),activity.u$steps)
activity.imp<-rbind(activity.imp,activity.u)
}
after<-activity.u$steps[1:20]#after imputation
#show before and after
cbind(before,after)[1:10,]

```

### Histogram for the imputed data, mean and median
The following histogram shows the total number of steps taken each day.

```{r}
daily.steps<-with(activity.imp,aggregate(steps,by=list(date),sum))
colnames(daily.steps) <-c("date","steps")
library(ggplot2)
ggplot(daily.steps, aes(x=steps)) +
    geom_histogram(fill="purple",binwidth=800) +
    labs(title="Histogram of daily steps",
         x = "Number of steps per day",y="Number of times in a day") + theme_bw()

mns2<-round(with(daily.steps, mean(steps,na.rm=TRUE)),2)
mdn2<-with(daily.steps, median(steps,na.rm=TRUE))
```
The mean total number of steps taken  per day after imputation is `r mns2`
The median total number of steps taken  per day is imputation is `r mdn2`

After imputing the missing data the estimated for mean has been reduced from `r mns1` to `r mns2` and similarly for median lowered from `r mdn1` to `r mdn2`

## Are there differences for activity between weekedays and weekends?
As we see from the time series plot for the weekdays there are more activity steps with peaks above  100.
While for weekends we have one peak which is way off the maximum peaks during weekdays. This will indicates the maximum  activity occured in an inerval falling on a weekend, but after it the rest are generally much lower than for the weekdays.
```{r}
library(lattice)
activity$date <- as.Date(activity$date)
activity$wday<-weekdays(activity$date)
activity$wday <-ifelse(substr(activity$wday,1,1)=="S","weekend","weekday")
par(mfrow=c(1,2))
#subset and aggregate steps for weekdays
wk.day<- subset(activity,activity$wday=="weekend")
colnames(wk.day) <- c("steps","date","interval","wkday")
wk.day<-with(na.omit(wk.day),aggregate(steps,by=list(interval),mean))


#with(wk.day, plot(steps,interval,type="l",col="red",lty=1,ylab="Average daily steps",xlab="5-minute interval",main="Weekdays"))
#subset and aggregate steps for weekend
wkend<- subset(activity,activity$wday!="weekend")
colnames(wkend) <- c("steps","date","interval","wkday")
wkend<-with(na.omit(wkend),aggregate(steps,by=list(interval),mean))
wkd<-cbind("weekday",wk.day);wek<-cbind("weekend",wkend)
colnames(wkd)<-c("day","steps","interval")
colnames(wek)<-c("day","steps","interval")


with(rbind(wkd,wek),xyplot(interval~steps|day,type="l"))
```

